\Title

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

    	Representation of the weights for $8$ directions for one speed (among the $8 \times 3$ different kernels of the model) as learned on the dataset of naturalistic scenes. Directions are shown as red arrows on the left. The spatio-temporal kernels are represented as slices of spatial weights at different delays. Delays varry along the horizontal axis from the far right (delay of one step) to the left (delay of $12$ steps). Each image corresponds to the weights at a given delay, with excitatory and inhibitory weights in warm and cold colors, respectively. Due to the observed symmetry between the ON and OFF event streams, we report that the kernels for the OFF polarities are very similar and are not shown here.	Different kernels are selective to the different motion directions and we observe a slight level of orientation selectivity perpendicular to the orientation. 

@@

If one focuses on the interpretation of these kernels in terms of spatio-temporal motifs embedded in the event stream, it can lead to interesting outcomes. In~\citep{grimaldi_robust_2022}, a link between event-based MLR training and Hebbian learning is drawn, allowing to say that the present model will learn its weights according to a presynaptic activity associated to the different motion directions. Each neuron becomes selective to a specific motion direction through the learning of an associated prototypical spatio-temporal spike motif. Each voxel in the 3D kernels defines a specific timestamp and a specific address. Consequently, our model is able to detect precise spatio-temporal motifs embedded in the spike train and associated to the different motion directions. The cone shape for the positive weights distribution highlights a loss of precision for longer delays, i.e. events away in the past. For the directions not coherent to the class of a training sample, an anti-Hebbian learning is also observed through the negative weights in the kernels of Fig.~\ref{fig:kernels}.

@@

%%%-----------------------------------------------------------------
%
\subsection{Testing with natural-like textures}
%%%-----------------------------------------------------------------
To test our model, we will quantify its ability to categorize different motions, that is on event streams for which the ground truth motion is known at every instant in time. Before applying the model on natural images, we will first test the model on simpler, parameterized stimuli. In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motion_clouds})~\citep{vacher_bayesian_2018}. In particular, we will set the spatial extent and duration of the generated movies similarly to the motion task defined above. This procedure defines a set of textures with different spatial properties and different motions picked from the same set of $8$ directions and $3$ speeds. For any given velocity, we also varied the parameters of the textures, such as the mean and variance of the orientation or spatial frequency content to provide with some naturalistic variability. This method provides a rich dataset of textured movies for which we know the ground truth for motion.
%%%-----------------------------------------------------------------
%: fig:motion_clouds
%%%-----------------------------------------------------------------
\begin{figure}%[h!]
    \centering
    \includegraphics[width=0.99\linewidth]{figures/motion_clouds.pdf} %
    \caption{{\bf Role of stimulus parameters in the motion detection accuracy.} Accuracy as a function of (from left to right) the mean spatial frequency, the bandwidth in spatial frequency (from a grating-like (left) to isotropic textures (right)), the bandwidth in orientation (from isotropic textures (left) to grating-like (right)), the bandwidth in speed (from a rigid motion (left) to independent frames (right)). Note that these accuracy is computed both in the case where orientation of the synthetic texture is necessarily perpendicular to the motion ('no aperture' condition) or in the generic case where orientation is independent of direction ('aperture').}
    \label{fig:motion_clouds}
\end{figure}
%%%-----------------------------------------------------------------

We plot here main axis of interests. First, as we change the mean spatial frequency of the texture, we observe a monotonous decrease in accuracy. This comes as a similar trend as that observed in the primary visual areas~\citep{priebe_tuning_2006} and revels the most informative scales learned by our model. We observe the tendency for different spatial frequency bandwidth. Notably, the accuracy is better for a grating-like stimulation than for a large (which qualitatively resemble a more textured stimulus), reminiscent to the behavioral response of humans to such stimuli~\citep{simoncini_more_2012}. Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth. When the stimulus is grating-like and that the orientation is arbitrary with respect to the direction of motion, the system is faced with the aperture problem and see a decrease of accuracy. This is not the case for isotropic stimuli or when the orientation is perpendicular to the direction of motion. Finally, we manipulated the amount of change between two successive frames, similar to a temperature parameter. This shows a progressive decrease in accuracy, similar to that observed in the amplitude of humans' eye movements~\citep{mansour_pour_speed_2018} but also that accuracy is low for a rigid motion which lacks variability.
%
\subsection{Accuracy efficiency trade-off}% for the motion detection task}
%%%-----------------------------------------------------------------
Once our MLR is trained, we obtain spatio-temporal kernels corresponding to the weights associated to the heterogeneous delays of our layer of spiking neurons and which may be used for detection. We observed that the distribution of the kernels' weights is sparse, with most values near zero (\seeFig.~\ref{fig:kernels}). As shown in the formalization of our event-based model, the computational cost of our model if implemented on a neuromorphic chip would be dominated by the number of spikes times the number of synapses. Indeed, the computations are dominated by the convolution operation. In a dense setting, this corresponds for all voxels in the output to a sum over all voxels in the inputs for all weights in the kernel. If the support of information is sparse, then computations can be performed only on those events. Also, if we set some weights of the kernels to zero, then the sum can be skipped for those addresses. Knowing the sparseness of the input, the total number of computations thus scales with the number of nonzero synaptic weights.

>>>>

    	Representation of the weights for $8$ directions for one speed (among the $8 \times 3$ different kernels of the model) as learned on the dataset of naturalistic scenes. Directions are shown as red arrows on the left. The spatio-temporal kernels are represented as slices of spatial weights at different delays. Delays vary along the horizontal axis from the far right (delay of one step) to the left (delay of $12$ steps). Each image corresponds to the weights at a given delay, with excitatory and inhibitory weights in warm and cold colors, respectively. Due to the observed symmetry between the ON and OFF event streams, we report that the kernels for the OFF polarities are very similar and are not shown here.	Different kernels are selective to the different motion directions and we observe a slight level of orientation selectivity perpendicular to the orientation. 

@@

If one focuses on the interpretation of these kernels in terms of spatio-temporal motifs embedded in the event stream, it can lead to interesting outcomes. In~\citep{grimaldi_robust_2022}, a link between event-based MLR training and Hebbian learning is drawn, allowing to say that the present model will learn its weights according to a presynaptic activity associated to the different motion directions. Each neuron becomes selective to a specific motion direction through the learning of an associated prototypical spatio-temporal spiking motif. Each voxel in the 3D kernels defines a specific timestamp and a specific address. Consequently, our model is able to detect precise spatio-temporal motifs embedded in the spike train and associated to the different motion directions. The cone shape for the positive weights' distribution highlights a loss of precision for longer delays, i.e. events away in the past. For the directions not coherent to the class of a training sample, an anti-Hebbian learning is also observed through the negative weights in the kernels of Fig.~\ref{fig:kernels}.

@@

%%%-----------------------------------------------------------------
%
\subsection{Testing with natural-like textures}
%%%-----------------------------------------------------------------
%%%-----------------------------------------------------------------
%: fig:motion_clouds
%%%-----------------------------------------------------------------
\begin{figure}%[h!]
    \centering
    \includegraphics[width=0.99\linewidth]{figures/motion_clouds.pdf} %
    \caption{{\bf Role of stimulus parameters in the motion detection accuracy.} Accuracy as a function of (from left to right) the mean spatial frequency, the bandwidth in spatial frequency (from a grating-like (left) to isotropic textures (right)), the bandwidth in orientation (from isotropic textures (left) to grating-like (right)), the bandwidth in speed (from a rigid motion (left) to independent frames (right)). Note that these accuracies are computed both in the case where orientation of the synthetic texture is necessarily perpendicular to the motion ('no aperture' condition) or in the generic case where orientation is independent of direction ('aperture').}
    \label{fig:motion_clouds}
\end{figure}
%%%-----------------------------------------------------------------
To test our model, we will quantify its ability to categorize different motions, that is on event streams for which the ground truth motion is known at every instant in time. Before applying the model on natural images, we will first test the model on simpler, parameterized stimuli. In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motion_clouds})~\citep{vacher_bayesian_2018}. In particular, we will set the spatial extent and duration of the generated movies similarly to the motion task defined above. This procedure defines a set of textures with different spatial properties and different motions picked from the same set of $8$ directions and $3$ speeds. For any given velocity, we also varied the parameters of the textures, such as the mean and variance of the orientation or spatial frequency content to provide with some naturalistic variability. This method provides a rich dataset of textured movies for which we know the ground truth for motion.

We plot here main axis of interests. First, as we change the mean spatial frequency of the texture, we observe a monotonous decrease in accuracy. This comes as a similar trend as that observed in the primary visual areas~\citep{priebe_tuning_2006} and revels the most informative scales learned by our model. We observe the tendency for different spatial frequency bandwidth. Notably, the accuracy is better for a grating-like stimulation than for a large (which qualitatively resemble a more textured stimulus), reminiscent to the behavioral response of humans to such stimuli~\citep{simoncini_more_2012}. Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth. When the stimulus is grating-like and that the orientation is arbitrary with respect to the direction of motion, the system is faced with the aperture problem and see a decrease of accuracy. This is not the case for isotropic stimuli or when the orientation is perpendicular to the direction of motion. Finally, we manipulated the amount of change between two successive frames, similar to a temperature parameter. This shows a progressive decrease in accuracy, similar to that observed in the amplitude of humans' eye movements~\citep{mansour_pour_speed_2018} but also that accuracy is low for a rigid motion which lacks variability.
%
\subsection{Accuracy efficiency trade-off}% for the motion detection task}
%%%-----------------------------------------------------------------
Once our MLR is trained, we obtain spatio-temporal kernels corresponding to the weights associated to the heterogeneous delays of our layer of spiking neurons and which may be used for detection. We observed that the distribution of the kernels' weights is sparse, with most values near zero (\seeFig{kernels}). As shown in the formalization of our event-based model, the computational cost of our model if implemented on a neuromorphic chip would be dominated by the computations used for the convolution operation. In a dense setting, this corresponds for all voxels in the output to a sum over all voxels in the inputs for all weights in the kernel. If the support of information is sparse, then computations can be performed only on those events. Also, if we set some weights of the kernels to zero, then the sum can be skipped for those addresses. Knowing the sparseness of the input, the total number of computations thus scales with the number of spikes multiplied by the number of nonzero synaptic weights. % , which is proportional to the number of spikes multiplied by the number of synapses
